---
title: "Bootstrap TC ~ TC pgs in PRAD"
author: "Nathalie Nilsson"
date: "November 19, 2019"
output: 
  html_document: 
    df_print: kable
    theme: paper
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

v3: New phenotype data used, compiled for baseline, young cases excluded

### Project
```{r}
path.to.project <- paste(rprojroot::find_rstudio_root_file(), "/", sep = "")
path.to.project
```

### Working directory relative to project
```{r}
gsub(rprojroot::find_rstudio_root_file(), "", getwd())
```


### Input
Input relative to project directory:
```{r}
score.files <- "pg_scoring/preventad/prad_v1/score_files/"
score.range <- "pg_scoring/preventad/prad_v1/scoring_range.txt"

dep.var <- "tc.mM"
covar.file <- "data/preventad/prad_covar_bl_data_v2.txt"
covars <- c("sex", "tc.age", "apoe4.bin")
int.var <- "statin"

output.name.org 	<- "boot_tc_v3"
output.dir      <- "analyses/preventad/boot_tc_v3/"
```

Input for analysis:
```{r}
n.pcs <- 5 # number of genetic principal components to include as covariates
n.bootstrap <- 100
```

### Libraries
```{r}
library(tidyverse)
library(rcompanion)
library(boot)
library(data.table)
```

### Data
```{r}
covar <- read.table(paste(path.to.project, covar.file, sep = "/"), 
                    header = TRUE, sep = "\t")
head(covar)

pcs <- paste("PC", 1:n.pcs, sep = "")
```

**Filter for non-missing values in tc.mM column**
```{r}
m.f <- filter(covar, !is.na(tc.mM)) %>%
  select("iid", pcs, covars, int.var, dep.var)
```

**Complete cases**
```{r}
m.all <- m.f[complete.cases(m.f), ] %>%
  rename(IID = iid)

head(m.all)
length(unique(m.all$IID)) == nrow(m.all)
```

**Excluded data**
```{r}
m.excl <- select(covar, iid, pcs, covars, int.var, dep.var) %>%
    filter(!iid %in% m.all$IID)
m.excl
```

**Descriptives**
```{r}
m.dt <- data.table(m.all)

# Stats ####
age.ttest <- t.test(m.dt[statin == 0, "tc.age"], m.dt[statin == 1, "tc.age"])
tc.ttest <- t.test(m.dt[statin == 0, "tc.mM"], m.dt[statin == 1, "tc.mM"])
apoe4.chisq <- chisq.test(table(m.all$statin, m.all$apoe4.bin))
sex.chisq <- chisq.test(table(m.all$statin, m.all$sex))

# Summarized data ####
m.g <- group_by(m.all, statin) %>%
    summarise(
        N_tot = length(IID),
        N_Females = sum(sex == "Female"), 
        Perc_Females = N_Females/N_tot,
        Perc_e4 = sum(apoe4.bin)/N_tot,
        Mean.age = mean(tc.age), 
        Mean.TC = mean(tc.mM)) %>%
    t() %>%
    data.frame()
m.g$p.value <- c(NA, NA, sex.chisq$p.value, sex.chisq$p.value, apoe4.chisq$p.value, age.ttest$p.value, tc.ttest$p.value)

write.table(m.g, paste(output.name.org, "_desc.txt", sep = ""), 
            col.names = FALSE, quote = FALSE, sep = "\t")

knitr::kable(apply(m.g, 2, function(x) round(x, 3)))
```


### PG scoring
```{r}
# New complete paths ####
score.files <- paste(path.to.project, score.files, sep = "/")
score.range <- paste(path.to.project, score.range, sep = "/")

output.dir      <- paste(path.to.project, output.dir, sep = "")

my.ls1 <- c(ls(), "my.ls1")
my.ls1
```

**All subjects**
```{r}
m.cc <- m.all
output.name <- output.name.org
covars <- c(covars[1:2], int.var)

source(paste(path.to.project, "scripts/lm_boot_forRmd_v2.R", sep = ""))

# Number of SNPs per score
n.snps <- select(m.cc, contains("CNTS")) %>%
    unique() %>%
    `row.names<-`("n.alleles") %>%
    t() %>%
    as.data.frame(optional = TRUE) %>%
    rownames_to_column() %>%
    separate(rowname, c("throw", "label"), sep = -3) %>%
    mutate(n.snps = n.alleles/2) %>%
    merge.data.frame(my.range.list)

n.snps
```

**Statin negative subjects**
```{r}
ls2 <- ls()
rm(list = ls2[!ls2 %in% my.ls1])

m.cc <- filter(m.all, statin == 0)
output.name <- paste(output.name.org, "_offstatin", sep = "")
covars <- covars[!grepl(int.var, covars)]

source(paste(path.to.project, "scripts/lm_boot_forRmd.R", sep = ""))
```

**Statin positive subjects**
```{r}
ls2 <- ls()
rm(list = ls2[!ls2 %in% my.ls1])

m.cc <- filter(m.all, statin == 1)
output.name <- paste(output.name.org, "_onstatin", sep = "")
covars <- covars[!grepl(int.var, covars)]

source(paste(path.to.project, "scripts/lm_boot_forRmd.R", sep = ""))
```

### Results plotted
```{r fig1, fig.height = 6, fig.width = 12, echo=FALSE}
all <- read.table(paste(output.name.org, "_Results.txt", sep = ""), header = TRUE) %>%
  mutate(pop = "all")
neg <- read.table(paste(output.name.org, "_offstatin_Results.txt", sep = ""), header = TRUE)  %>%
  mutate(pop = "Off statins")
pos <- read.table(paste(output.name.org, "_onstatin_Results.txt", sep = ""), header = TRUE) %>%
  mutate(pop = "On statins")

m2 <- rbind.data.frame(all, neg, pos) %>%
  merge.data.frame(my.range.list)

ggplot(m2, aes(x = factor(pmax), y = Mean.value.R2, color = -log10(Original.value.P))) +
    geom_point(size = 3, shape = 18) +
    geom_errorbar(aes(ymin = Mean.value.R2 - SD.value.R2, ymax = Mean.value.R2 + SD.value.R2), 
                  width = 0.3, size = 1) +
    labs(title = paste(dep.var, "~ PGS"), 
         caption = paste("corrected for:", paste(c(paste("PC1-", n.pcs, sep = ""), covars, "and statin use for the 'all' group"), collapse = ", "), "\nerror bars indicate SD"), 
         y = "Additional R2 explained by PGS", 
         x = "P cut-off") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   facet_grid(~pop)
```

### TC ~ statin use * TC-pgs
```{r}
tc.data <- read.table(paste(output.name.org, "_data.txt", sep = ""), header = TRUE)
head(tc.data)
```

```{r}
ggplot(tc.data, aes(SCORES06, tc.mM, color = factor(statin))) +
  geom_point() +
  geom_smooth(method = "lm")
```

